---
title: "Project P2"
date: "2024-03-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data = read.csv("data- cleaned.csv")
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

```{r}
summary(data)
```

```{r}
data$gender = as.factor(data$gender)
data$ever_married = as.factor(data$ever_married)
data$work_type = as.factor(data$work_type)
data$Residence_type = as.factor(data$Residence_type)
data$smoking.status = as.factor(data$smoking.status)
data = na.omit(data)
data <- data[data$age >= 45, ]
data$work_type <- droplevels(data$work_type)

#a <- quantile(data$age,0.25)
#b <- quantile(data$age,0.75)
#IQR <- b-a
#lb <- a-1.5*IQR
#ub <- b-1.5*IQR
#data <- subset(data, data$age >= lb & data$age <= ub)
```

```{r}
summary(data)
```


Fit full model:
```{r}
modelfull <- glm(stroke~.-id-X, data = data, family = "binomial"(link = "logit"))
summary(modelfull)
```

AIC
```{r}
## Stepwise elimination based on AIC ##
sel.var.aic <- step(modelfull, trace = 0, k = 2, direction = "both") 
select_var_aic<-attr(terms(sel.var.aic), "term.labels")   
select_var_aic

```

BIC
```{r}
## Stepwise elimination based on AIC ##
sel.var.bic <- step(modelfull, trace = 0, k = log(nrow(data)), direction = "both") 
select_var_bic<-attr(terms(sel.var.bic), "term.labels")   
select_var_bic

```

AIC stepwise selection model1:

```{r}
model1 <- glm(stroke ~ age + hypertension + heart_disease + ever_married + work_type + avg_glucose_level, 
              data = data, family = "binomial"(link = "logit"))
summary(model1)
```

```{r}
e_hat <- resid(model1)
qqnorm(e_hat)
qqline(e_hat)
```

BIC stepwise selection model2:

```{r}
model2 <- glm(stroke ~ age + avg_glucose_level, data = data, family = "binomial"(link = "logit"))
summary(model2)
```

```{r}
model <- lm(stroke ~ .-X-id, data = data)
conf_intervals <- confint(model)
print(conf_intervals)

```

```{r}
anova(modelfull, test = 'Chisq')

```

```{r}
anova(sel.var.aic, test = 'Chisq')

```

p-value is good and significant.


```{r}
anova(sel.var.bic, test = 'Chisq')

```

LASSO
```{r}
library(glmnet)
library(survival)
library(MASS)
library(influence.ME)
x = as.matrix(data[,c(3:10,13)])
y = data$stroke
fit = glmnet(x, y, family = "binomial")
plot(fit, xvar = "dev", label = TRUE)
cv.out = cv.glmnet(x, y, family = "binomial", type.measure = "class", alpha = 0.5)
plot(cv.out)
best.lambda <- cv.out$lambda.1se
best.lambda
co <- coef(cv.out, s = "lambda.1se")
co

```
LASSO not good

```{r}
vif(model1)
```

VIF is good except work_type, all around 1.

```{r}
install.packages("rms")
library(rms)
```

```{r, echo=FALSE, message=FALSE, results='hide', fig.width=11, fig.height=5}
par(mfrow = c(1, 2))
set.seed(1008088070)
lrm.final <- lrm(stroke~., data = data[,which(colnames(data) %in% c("age","hypertension", 
                                                                    "heart_disease", "ever_married", "work_type",
                                                                    "avg_glucose_level", "stroke"))], 
                 x =TRUE, y = TRUE, model= T)
cross.calib <- calibrate(lrm.final, method="crossvalidation", B=10) # model calibration
plot(cross.calib, las=1, xlab = "Predicted Probability", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.7, cex.sub = 0.7)

library(pROC)
p <- predict(lrm.final1, type = "fitted")

roc_logit <- roc(data$stroke ~ p)
## The True Positive Rate ##
TPR <- roc_logit$sensitivities
## The False Positive Rate ##
FPR <- 1 - roc_logit$specificities

plot(cex.main = 0.8, cex.lab = 0.8, FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2, col = 'red')
abline(a = 0, b = 1, lty = 2, col = 'blue')
text(0.7,0.4,label = paste("AUC = ", round(auc(roc_logit),2)))
auc(roc_logit)
```

```{r, echo=FALSE, message=FALSE, results='hide', fig.width=11, fig.height=5}
par(mfrow = c(1, 2))
set.seed(1008088070)
lrm.final <- lrm(stroke~., data = data[,which(colnames(data) %in% c("age",
                                                                    "avg_glucose_level", "stroke"))], 
                 x =TRUE, y = TRUE, model= T)
cross.calib <- calibrate(lrm.final, method="crossvalidation", B=10) # model calibration
plot(cross.calib, las=1, xlab = "Predicted Probability", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.7, cex.sub = 0.7)

library(pROC)
p <- predict(lrm.final, type = "fitted")

roc_logit <- roc(data$stroke ~ p)
## The True Positive Rate ##
TPR <- roc_logit$sensitivities
## The False Positive Rate ##
FPR <- 1 - roc_logit$specificities

plot(cex.main = 0.8, cex.lab = 0.8, FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2, col = 'red')
abline(a = 0, b = 1, lty = 2, col = 'blue')
text(0.7,0.4,label = paste("AUC = ", round(auc(roc_logit),2)))
auc(roc_logit)
```


```{r}
## Fit the model with lrm from rms package ##
set.seed(1008088070)
lrm.final <- lrm(stroke ~ ., data = data[,which(colnames(data) %in% c(select_var_aic, "stroke"))], x =TRUE, y = TRUE, model= T)
cross.calib <- calibrate(lrm.final, method="crossvalidation", B=10) # model calibration
plot(cross.calib, las=1, xlab = "Predicted Probability")

```

Limitaton


library(pROC)
p <- predict(lrm.final, type = "fitted")

roc_logit <- roc(data$stroke ~ p)
## The True Positive Rate ##
TPR <- roc_logit$sensitivities
## The False Positive Rate ##
FPR <- 1 - roc_logit$specificities

plot(FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2, col = 'red')
abline(a = 0, b = 1, lty = 2, col = 'blue')
text(0.7,0.4,label = paste("AUC = ", round(auc(roc_logit),2)))

auc(roc_logit)



AUC is good

```{r}
#residuals <- residuals(model1)
#eta <- model1
#plot(eta, residuals, xlab = "Eta (x*beta)", ylab = "Residuals", main = "QQ-Plot")
#abline(h = 0, col = "red")  # Add a horizontal line at y = 0 for reference)
```




```{r}
## Dfbetas ##
par(mfrow = c(1, 2))

log.mod.final <- glm(stroke ~ age + hypertension + heart_disease + ever_married + work_type + 
                       avg_glucose_level, 
                     family = binomial, data = data)
df.final <- dfbetas(log.mod.final)
head(df.final)

par(family = 'serif')
plot(data$age, df.final[,1], xlab='age of patients', 
     ylab='dfbeta')
lines(lowess(data$age, df.final[,1]), lwd=2, col='blue')
abline(h=0, lty='dotted')
abline(h=-2/sqrt(nrow(df.final)), lty='dotted')
abline(h=2/sqrt(nrow(df.final)), lty='dotted')


log.mod.final3 <- glm(stroke ~  age + hypertension + heart_disease + ever_married + work_type + 
                       avg_glucose_level,
                     family = binomial, data = data)
df.final3 <- dfbetas(log.mod.final3)
head(df.final3)

par(family = 'serif')
plot(data$avg_glucose_level, df.final3[,5], xlab='average glucose level of patients', 
     ylab='dfbeta')
lines(lowess(data$avg_glucose_level, df.final3[,5]), lwd=2, col='blue')
abline(h=0, lty='dotted')
abline(h=-2/sqrt(nrow(df.final3)), lty='dotted')
abline(h=2/sqrt(nrow(df.final3)), lty='dotted')

```

```{r}
par(mfrow = c(1, 2))
log.mod.final1 <- glm(stroke ~ age + hypertension + heart_disease + ever_married + work_type + 
                       avg_glucose_level,
                      family = binomial, data = data)
df.final1 <- dfbetas(log.mod.final1)

# Assuming 'hypertension' is a factor/categorical variable
par(family = 'serif')
boxplot(df.final1[,2] ~ data$hypertension, xlab = 'Hypertension of patients', 
        ylab = 'DFBETA', main = 'DFBETA check for Hypertension', col = 'lightblue')

# Add horizontal lines for reference
abline(h = 0, lty = 'dotted')
abline(h = -2 / sqrt(nrow(df.final1)), lty = 'dotted')
abline(h = 2 / sqrt(nrow(df.final1)), lty = 'dotted')


log.mod.final2 <- glm(stroke ~ age + hypertension + heart_disease + ever_married + work_type + 
                       avg_glucose_level,
                      family = binomial, data = data)
df.final2 <- dfbetas(log.mod.final2)

# Assuming 'hypertension' is a factor/categorical variable
par(family = 'serif')
boxplot(df.final2[,3] ~ data$hypertension, xlab = 'Heart disease of patients', 
        ylab = 'DFBETA', main = 'DFBETA check for Heart disease', col = 'lightblue')

# Add horizontal lines for reference
abline(h = 0, lty = 'dotted')
abline(h = -2 / sqrt(nrow(df.final2)), lty = 'dotted')
abline(h = 2 / sqrt(nrow(df.final2)), lty = 'dotted')
```


dfbeta for average glucose level of patients is good. not pattern and all around 0. 
There may exist some influential points for average glucose level of patients data.


New method: 
```{r}
D_cut <- 4/2577
cook_dist <- cooks.distance(model1)
plot(cook_dist)
influential1 <- which(abs(cook_dist) > D_cut)
```

```{r}
D_cut2 <- 4/2577
cook_dist2 <- cooks.distance(model2)
plot(cook_dist2)
influential2 <- which(abs(cook_dist2) > D_cut2)
```

```{r}
fits_cut <- 2*sqrt((4+1)/2577)
dffits_i <- dffits(model1)
plot(dffits_i)
influential3 <- which(abs(dffits_i) > fits_cut)
```

```{r}
fits_cut4 <- 2*sqrt((2+1)/2577)
dffits_i4 <- dffits(model2)
plot(dffits_i4)
influential4 <- which(abs(dffits_i4) > fits_cut4)

```



```{r}
cleaned_data1 <- data[-influential1, ]
```



```{r}
modelfull1 <- glm(stroke~.-id-X, data = cleaned_data1, family = "binomial"(link = "logit"))
summary(modelfull1)

```

```{r}
cleaned_data2 <- data[-influential2, ]

```

```{r}
modelfull2 <- glm(stroke~.-id-X, data = cleaned_data2, family = "binomial"(link = "logit"))
summary(modelfull2)

```

```{r}

D_cut <- 4/2577
cook_dist <- cooks.distance(model3)
plot(cook_dist)
#influential1 <- which(abs(cook_dist) > D_cut)
```


```{r}
## Stepwise elimination based on AIC ##
sel.var.aic1 <- step(modelfull1, trace = 0, k = 2, direction = "both") 
select_var_aic1<-attr(terms(sel.var.aic1), "term.labels")   
select_var_aic1

```



```{r}
model3 <- glm(stroke ~ gender + age + hypertension + heart_disease + work_type + smoking.status, 
              data = cleaned_data1, family = "binomial"(link = "logit"))
summary(model3)

```

Not good enough, the standard error is large and the variables are not siginicant. 
So still use the original AIC model.

Final model: 
```{r}
model4 <- glm(stroke ~ age + hypertension + heart_disease + avg_glucose_level, 
              data = data, family = "binomial"(link = "logit"))
summary(model4)

```

```{r, echo=FALSE, message=FALSE, results='hide', fig.width=11, fig.height=5}
par(mfrow = c(1, 2))
set.seed(1008088070)
lrm.final <- lrm(stroke~., data = data[,which(colnames(data) %in% c("age","hypertension", 
                                                                    "heart_disease",
                                                                    "avg_glucose_level", "stroke"))], 
                 x =TRUE, y = TRUE, model= T)
cross.calib <- calibrate(lrm.final, method="crossvalidation", B=10) # model calibration
plot(cross.calib, las=1, xlab = "Predicted Probability", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.7, cex.sub = 0.7)

library(pROC)
p <- predict(lrm.final, type = "fitted")

roc_logit <- roc(data$stroke ~ p)
## The True Positive Rate ##
TPR <- roc_logit$sensitivities
## The False Positive Rate ##
FPR <- 1 - roc_logit$specificities

plot(cex.main = 0.8, cex.lab = 0.8, FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2, col = 'red')
abline(a = 0, b = 1, lty = 2, col = 'blue')
text(0.7,0.4,label = paste("AUC = ", round(auc(roc_logit),2)))
auc(roc_logit)

```





```{r}
sel.var.bic1 <- step(modelfull2, trace = 0, k = log(nrow(data)), direction = "both") 
select_var_bic1<-attr(terms(sel.var.bic1), "term.labels")   
select_var_bic1
```
not applicable, so we need to choose the original model.


# EDA
```{r}
hypercounts <- data %>%
  count(hypertension) %>%
  mutate(Hypertension = ifelse(hypertension == 0, "No", "Yes"))

custom_colors <- c("No" = "red",
                   "Yes" = "blue")
ggplot(hypercounts, aes(x = Hypertension, y = n, fill = Hypertension)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), vjust = -0.5) +
  labs(title = "Hypertension of Patients", x = "Hypertension", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none") +
  scale_fill_manual(values = custom_colors)
```


```{r}
heartcounts <- data %>%
  count(heart_disease) %>%
  mutate(Heart_disease = ifelse(heart_disease == 0, "No", "Yes"))

custom_colors <- c("No" = "red",
                   "Yes" = "blue")
ggplot(heartcounts, aes(x = Heart_disease, y = n, fill = Heart_disease)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), vjust = -0.5) +
  labs(title = "Heart disease of Patients", x = "Heart disease", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none") +
  scale_fill_manual(values = custom_colors)

```


```{r}
strokecount <- data %>%
  count(stroke) %>%
  mutate(Stroke = ifelse(stroke == 0, "No", "Yes"))

custom_colors <- c("No" = "red",
                   "Yes" = "blue")
ggplot(strokecount, aes(x = Stroke, y = n, fill = Stroke)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), vjust = -0.5) +
  labs(title = "Stroke of Patients", x = "Stroke", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none") +
  scale_fill_manual(values = custom_colors)

```

```{r}
boxplot(data$age,main="Boxplot of Patient Age",ylab="Age",las=1)
```

```{r}
boxplot(data$avg_glucose_level, main="Boxplot of average glucose level", ylab="average glucose", las=1)
```

```{r}
boxplot(data$bmi, main="Boxplot of BMI level", ylab="BMI", las=1)
```

```{r}
data$stroke_status <- factor(data$stroke, levels = c(0, 1), labels = c("No Stroke", "Stroke"))

boxplot(data$avg_glucose_level[data$stroke_status == "No Stroke"], 
        data$avg_glucose_level[data$stroke_status == "Stroke"], 
        main = "Boxplot of average glucose level by Stroke Status", 
        ylab = "average glucose level", 
        names = c("No Stroke", "Stroke"),
        las = 1)
```

```{r}
data$stroke_status <- factor(data$stroke, levels = c(0, 1), labels = c("No Stroke", "Stroke"))

boxplot(data$age[data$stroke_status == "No Stroke"], 
        data$age[data$stroke_status == "Stroke"], 
        main = "Boxplot of age  by Stroke Status", 
        ylab = "age", 
        names = c("No Stroke", "Stroke"),
        las = 1)
```



```{r}
data$stroke_status <- factor(data$stroke, levels = c(0, 1), labels = c("No Stroke", "Stroke"))

boxplot(data$bmi[data$stroke_status == "No Stroke"], 
        data$bmi[data$stroke_status == "Stroke"], 
        main = "Boxplot of BMI level by Stroke Status", 
        ylab = "BMI", 
        names = c("No Stroke", "Stroke"),
        las = 1)
```

```{r}
worktypecount <- data %>%
  count(work_type) %>%
  mutate(Work_type = work_type, Frequency = n)

ggplot(worktypecount, aes(x = Work_type, y = Frequency, fill = Work_type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = 0) +
  labs(title = "Work type of Patients", x = "Work type", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")

```

```{r}
restypecount <- data %>%
  count(Residence_type) %>%
  mutate(Residence_Type = Residence_type, Frequency = n)

ggplot(restypecount, aes(x = Residence_Type, y = Frequency, fill = Residence_Type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = 0) +
  labs(title = "Residence type of Patients", x = "Residence type", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")

```

```{r}
gendertypecount <- data %>%
  count(gender) %>%
  mutate(gender = gender, Frequency = n)

ggplot(gendertypecount, aes(x = gender, y = Frequency, fill = gender)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = 0) +
  labs(title = "Gender of Patients", x = "Gender", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")

```

```{r}
marriedcount <- data %>%
  count(ever_married) %>%
  mutate(ever_married = ever_married, Frequency = n)

ggplot(marriedcount, aes(x = ever_married, y = Frequency, fill = ever_married)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = 0) +
  labs(title = "Marital status of Patients", x = "Marital status", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")

```











